### 2024-03-11 paper/paper.qmd
Draft of codes

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Yanyu Wu
  - Ziyi Liu
  - Hechen Zhang
thanks: "Code and data are available at: https://github.com/HechenZ123/Cause-of-Deaths-in-Alberta.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(broom.mixed)
library(tidyverse)
library(janitor)
library(knitr)
library(modelsummary)
library(rstanarm)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
alberta_cod <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
alberta_cod |>
  filter(
    calendar_year == 2022,
    ranking <= 8
  ) |>
  mutate(total_deaths = format(total_deaths, big.mark = ",")) |>
  kable(
    col.names = c("Year", "Cause", "Ranking", "Deaths", "Years"),
    align = c("l", "r", "r", "r", "r"),
    digits = 0, booktabs = TRUE, linesep = ""
  )
```
```{r}
# Look up the five most common causes of death in 2022 of those that have been present every year.
alberta_cod_top_five <-
  alberta_cod |>
  filter(
     calendar_year == 2022
  ) |>
  slice_max(order_by = desc(ranking), n = 5) |>
  pull(cause)
head(alberta_cod_top_five)

alberta_cod <-
  alberta_cod |>
  filter(cause %in% alberta_cod_top_five)
head(alberta_cod)
```

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false
alberta_cod |>
  ggplot(aes(x = calendar_year, y = total_deaths, color = cause)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Year", y = "Annual number of deaths in Alberta") +
  facet_wrap(vars(cause), dir = "v", ncol = 1) +
  theme(legend.position = "none")
```

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

cause_of_death_alberta_poisson <-
  readRDS(file = here::here("models/cause_of_death_alberta_poisson.rds"))

cause_of_death_alberta_neg_binomial <-
  readRDS(file = here::here("models/cause_of_death_alberta_neg_binomial.rds"))
```

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
coef_short_names <- 
  c("causeOrganic dementia"
    = "causeOrganic dementia",
    "causeOther ill-defined and unkno..."
    = "causeOther ill-defined...",
    "causeCOVID-19, virus identified"
    = "causeCOVID-19",
    "causeMalignant neoplasms of trac..."
    = "causeMalignant neoplasms of trac..."
    )

modelsummary(
  list(
    "Poisson" = cause_of_death_alberta_poisson,
    "Negative binomial" = cause_of_death_alberta_neg_binomial
  ),
  coef_map = coef_short_names
)

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(cause_of_death_alberta_poisson) +
  theme(legend.position = "bottom")

pp_check(cause_of_death_alberta_neg_binomial) +
  theme(legend.position = "bottom")
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

poisson <- loo(cause_of_death_alberta_poisson, cores = 2)
neg_binomial <- loo(cause_of_death_alberta_neg_binomial, cores = 2)

loo_compare(poisson, neg_binomial)
```

\newpage

# References